A Horrocks-type theorem for even orthogonal $K_2$

### 2020-04-09 dl-paper.tex
Cosmetics

\begin{corollary} \label{K-a-b-cor2}
 One has $\widetilde{P}_{\alpha, M}(0) \cap K(A[X], M[X]) = \widetilde{K}(\alpha, \beta) \cdot X_{-\beta}(M \cdot X).$ 
\end{corollary}
\begin{proof} Set $L = \widetilde{P}_{\alpha, M}(0) \cap K(A[X], M[X])$.
The elements of $X_{-\beta}(M\cdot X)$ are generators of type P1 for $\widetilde{P}_{\alpha, M}(0)$ (cf.~\cref{rem:recognition}),
 therefore $X_{-\beta}(M \cdot X) \subseteq L$. On the other hand, by~\cref{K-a-b-cor} $\widetilde{K}(\alpha, \beta) \subseteq L$. Thus, we have shown that $\widetilde{K}(\alpha, \beta) \cdot X_{-\beta}(M \cdot X) \subseteq L$. The reverse inclusion immediately follows from~\cref{K-a-b} and~\cref{rem:palpha}.
 \end{proof}

@@

We come to the central concept of this subsection.
\begin{df} Define the subgroups $P_\alpha(0)$, $P_{\alpha}(*)$ and $K(\alpha, \beta)$ of $\overline{\St}(\Phi, R, M[X, X^{-1}])$ as the images of the subgroups $\widetilde{P}_{\alpha, M}(0)$, $\widetilde{P}_{\alpha, M}(*)$ and $\widetilde{K}(\alpha, \beta)$ under the natural homomorphism $j_+ \colon \St(\Phi, A[X]) \to \St(\Phi, R)$.\end{df}

@@

\begin{corollary} \label{K-a-b-P-b} The image of $K(\alpha, \beta)$ under the automorphism of conjugation by $x_\alpha(aX^{-1})$ is contained in $P_\beta(0)$. \end{corollary}
\begin{proof} 
In view of~\cref{image-K-a-b} and~\cref{K-a-b-cor} it suffices to show that $Z_{\alpha, \beta} \subseteq P_\beta(0)$. Let $g = x_{\gamma}(f)$ be a generator of $Z_{\alpha, \beta}$ for some
$\gamma \in Z_+(\alpha) \setminus \{\alpha - \beta \}$. Notice that $\gamma$ cannot form an obtuse angle with $\beta$, otherwise from $\langle \beta, \gamma \rangle = -1$ it follows that $ \langle \alpha - \beta, \gamma \rangle \geq 2$ and hence that $\gamma = \alpha - \beta$, a contradiction.
Now if $\langle \beta, \gamma \rangle$ is $0$, $1$ or $2$, then $g$ is a generator of type P3, P2 or P5 for $P_\beta(0)$, respectively.\end{proof}

@@

\begin{lemma} \label{conj-S-a-x-bma} For $a, b\in A$ the image of $K(\alpha,\beta)^{x_\alpha(aX^{-1})}$ under the automorphism of conjugation by $x_{\beta-\alpha}(b)$ is contained in $P_\alpha(0)$. \end{lemma}
\begin{proof} By Lemmas~\ref{conj-K-a-b}--\ref{image-K-a-b} it suffices to show that the conjugate $g_1$ of each generator $g = x_\gamma(f)$, $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$ of $Z_{\alpha, \beta}$ by $x_{\beta-\alpha}(b)$ belongs to $P_\alpha(0)$. It is clear that $g$ itself is a generator of type P2 for $P_\alpha(0)$, therefore it remains to consider the case $\gamma\not \perp \beta - \alpha$.

@@

\begin{lemma} \label{SR:obtuse} For $g \in j_+(K(A[X], M[X])) \cap P_\alpha(m) \cap P_\beta(m')$ and $a, b\in A$ the element \[ x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}((1 + \epsilon abm')^{-1})\cdot x_{\beta - \alpha}(-b(1 + \epsilon abm')) \]
 belongs to $P_\alpha\left(\tfrac{m - \epsilon bm'}{1+am}\right)$, where $\epsilon = N_{\alpha, -\beta}$.
\end{lemma}
\begin{proof} Set $c = 1 + \epsilon a b m'$, $s_0 = \langle a, m \rangle$, $s_1 = \langle \epsilon am', -bc^{-1} \rangle$.
Notice that $1 - \epsilon abc^{-1}m' = c^{-1}$, therefore by~\eqref{eq:dennis-stein}
\begin{equation} \label{eq:computation-SR-obtuse} x_{\alpha-\beta}(\epsilon am') \cdot x_{\beta-\alpha}(-bc^{-1}) = x_{\beta-\alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot x_{\alpha-\beta}(\epsilon acm').\end{equation}
Notice that $x_{-\beta}(m'X) \in P_\alpha(0)$, therefore from the definition of $S_\alpha(a, -)$ we obtain that
\begin{align*}
 S_\alpha\left(a, x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)\right) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta-\alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) \\ 
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot h_\alpha^{-1} (1 + am) \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Smult}}\\
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Tulenbaev-formula}}\\
 = x_{-\beta}(m'X) \cdot x_{\alpha - \beta}(\epsilon am') \cdot x_{\beta - \alpha}(-bc^{-1}) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{Chevalley-CCF1},~\eqref{Chevalley-CCF2}, \eqref{eq:conj-h-x}}\\
 \begin{split}= x_{-\beta}(m'X) \cdot x_{\beta - \alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot \hspace{100pt} \\ \cdot x_{\alpha-\beta}\left(\epsilon acm'\right) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{eq:computation-SR-obtuse}} \end{split}\\
 =  x_{\beta - \alpha}(-b) \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(-\epsilon bm'X) \cdot s_1 \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot x_{-\alpha}\left(\tfrac{cmX}{1+am}\right) \cdot s_0 &\text{ by~\eqref{Chevalley-CCF1},~\eqref{eq:conj-h-x},~\eqref{eq:h-inv}}\\
 = x_{\beta-\alpha}(-b) \cdot x_{-\beta}(m'c^{-1}X) \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot s_1 \cdot s_0 \cdot x_{-\alpha}\left(-\epsilon bm' X + \tfrac{cmX}{1+am}\right) &\text{ by~\eqref{Chevalley-CCF1}.}
\end{align*}
Denote by $h_0$ the product of all factors in the above formula except the first and the last one. It is clear that $h_0$ lies in $P_\alpha(0)$. Using~\cref{K-a-b} we can decompose $g$ as $g_0 \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)$ for some $g_0 \in K(\alpha, \beta)$, $m, m' \in M$. The required assertion now follows from \cref{lem:Smult},\cref{conj-S-a-x-bma} and the above computation:
\begin{multline} \nonumber
 x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) = \\
  = x_{\beta - \alpha}(b) \cdot S_\alpha(a, g_0) \cdot x_{\beta-\alpha}(-b) \cdot h_0 \cdot x_{-\alpha}\left(\tfrac{m -\epsilon bm'}{1+am}\right) \in P_\alpha(\tfrac{m -\epsilon bm'}{1+am}). \qedhere
\end{multline}
\end{proof}

@@

\subsection{Proof of Horrocks theorem.} \label{sec:P1glueing}
We retain the notation of the previous section and assume additionally that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, R)$ is injective.
This assumption allows us to consider the group $\St(\Phi, B)$ as a subgroup of $\St(\Phi, R)$.
Being considered in this way, $\St(\Phi, B)$ contains the subgroup $\overline{G}^{\geq 0}_M$ of $\St(\Phi, R)$.
This follows from a consideration of the following commutative diagram, in which the map $t$ is obtained from~\cref{lem:lemma32}:
\[ \xymatrix{ \St(\Phi, A[X], M[X]) \ar[d] \ar[r] & \St(\Phi, R, M[X, X\inv]) \ar[r]^(.65){t} & \St(\Phi, B) \ar[d]^{j_R} \\
   \St(\Phi, A[X]) \ar[rr]^{j_+} & & \St(\Phi, R). } \]

>>>>

\begin{corollary} \label{K-a-b-cor2}
 One has $\widetilde{P}_{\alpha, M}(0) \cap K(A[X], M[X]) = \widetilde{K}(\alpha, \beta) \cdot X_{-\beta}(M \cdot X).$ 
\end{corollary}
\begin{proof} Set $L = \widetilde{P}_{\alpha, M}(0) \cap K(A[X], M[X])$.
The elements of $X_{-\beta}(M\cdot X)$ are generators of type P1 for $\widetilde{P}_{\alpha, M}(0)$ (cf.~\cref{rem:recognition}),
 therefore $X_{-\beta}(M \cdot X) \subseteq L$. On the other hand, by~\cref{K-a-b-cor} $\widetilde{K}(\alpha, \beta) \subseteq L$. Thus, we have shown that $\widetilde{K}(\alpha, \beta) \cdot X_{-\beta}(M \cdot X) \subseteq L$. The reverse inclusion immediately follows from~\cref{rem:palpha} and~\cref{K-a-b}.
 \end{proof}

@@

We come to the central concept of this subsection.
\begin{df} Define the subgroups $P_\alpha(0)$, $P_{\alpha}(*)$ and $K(\alpha, \beta)$ of $\overline{\St}(\Phi, R, I)$ as the images of the subgroups $\widetilde{P}_{\alpha, M}(0)$, $\widetilde{P}_{\alpha, M}(*)$ and $\widetilde{K}(\alpha, \beta)$ under the natural homomorphism $j_+ \colon \St(\Phi, A[X]) \to \St(\Phi, R)$.\end{df}

@@

\begin{corollary} \label{K-a-b-P-b} The image of $K(\alpha, \beta)$ under the automorphism of conjugation by $x_\alpha(aX^{-1})$ is contained in $P_\beta(0)$. \end{corollary}
\begin{proof} 
By~\cref{K-a-b-cor} and~\cref{image-K-a-b} it suffices to show that $Z_{\alpha, \beta} \subseteq P_\beta(0)$. Let $g = x_{\gamma}(f)$ be a generator of $Z_{\alpha, \beta}$ for some
$\gamma \in Z_+(\alpha) \setminus \{\alpha - \beta \}$. Notice that $\gamma$ cannot form an obtuse angle with $\beta$, otherwise from $\langle \beta, \gamma \rangle = -1$ it follows that $ \langle \alpha - \beta, \gamma \rangle \geq 2$ and hence that $\gamma = \alpha - \beta$, a contradiction.
Now if $\langle \beta, \gamma \rangle$ is $0$, $1$ or $2$, then $g$ is a generator of type P3, P2 or P5 for $P_\beta(0)$, respectively.\end{proof}

@@

\begin{lemma} \label{conj-S-a-x-bma} For $a, b\in A$ the image of $K(\alpha,\beta)^{x_\alpha(aX^{-1})}$ under the automorphism of conjugation by $x_{\beta-\alpha}(b)$ is contained in $P_\alpha(0)$. \end{lemma}
\begin{proof} By Lemmas~\ref{conj-K-a-b}, \ref{image-K-a-b} it suffices to show that the conjugate $g_1$ of each generator $g = x_\gamma(f)$, $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$ of $Z_{\alpha, \beta}$ by $x_{\beta-\alpha}(b)$ belongs to $P_\alpha(0)$. It is clear that $g$ itself is a generator of type P2 for $P_\alpha(0)$, therefore it remains to consider the case $\gamma\not \perp \beta - \alpha$.

@@

\begin{lemma} \label{SR:obtuse} For $g \in j_+(K(A[X], M[X])) \cap P_\alpha(m) \cap P_\beta(m')$ and $a, b\in A$ the element \[ x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}((1 + \epsilon abm')^{-1})\cdot x_{\beta - \alpha}(-b(1 + \epsilon abm')) \]
 belongs to $P_\alpha\left(\tfrac{m - \epsilon bm'}{1+am}\right)$, where $\epsilon = N_{\alpha, -\beta}$.
\end{lemma}
\begin{proof} Set $c = 1 + \epsilon a b m'$, $s_0 = \langle a, m \rangle$, $s_1 = \langle \epsilon am', -bc^{-1} \rangle$.
Notice that $1 - \epsilon abc^{-1}m' = c^{-1}$, therefore by~\eqref{eq:dennis-stein}
\begin{equation} \label{eq:computation-SR-obtuse} x_{\alpha-\beta}(\epsilon am') \cdot x_{\beta-\alpha}(-bc^{-1}) = x_{\beta-\alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot x_{\alpha-\beta}(\epsilon acm').\end{equation}
Notice that $x_{-\beta}(m'X) \in P_\alpha(0)$, therefore from the definition of $S_\alpha(a, -)$ we obtain that
\begin{align*}
 S_\alpha\left(a, x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)\right) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta-\alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) \\ 
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot S_\alpha(a, x_{-\alpha}(mX)) \cdot h_\alpha^{-1} (1 + am) \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Smult}}\\
 = S_\alpha(a, x_{-\beta}(m'X)) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) \cdot x_{\beta - \alpha}(-bc) &\text{ by~\cref{lem:Tulenbaev-formula}}\\
 = x_{-\beta}(m'X) \cdot x_{\alpha - \beta}(\epsilon am') \cdot x_{\beta - \alpha}(-bc^{-1}) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{Chevalley-CCF1},~\eqref{Chevalley-CCF2}, \eqref{eq:conj-h-x}}\\
 \begin{split}= x_{-\beta}(m'X) \cdot x_{\beta - \alpha}(-b) \cdot s_1 \cdot h_{\alpha-\beta}(c^{-1}) \cdot \hspace{100pt} \\ \cdot x_{\alpha-\beta}\left(\epsilon acm'\right) \cdot x_{-\alpha}\left(\tfrac{mX}{1+am}\right) \cdot s_0 \cdot h_{\beta - \alpha}(c^{-1}) &\text{ by~\eqref{eq:computation-SR-obtuse}} \end{split}\\
 =  x_{\beta - \alpha}(-b) \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(-\epsilon bm'X) \cdot s_1 \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot x_{-\alpha}\left(\tfrac{cmX}{1+am}\right) \cdot s_0 &\text{ by~\eqref{Chevalley-CCF1},~\eqref{eq:conj-h-x},~\eqref{eq:h-inv}}\\
 = x_{\beta-\alpha}(-b) \cdot x_{-\beta}(m'c^{-1}X) \cdot x_{\alpha-\beta}(\epsilon ac^{-1}m') \cdot s_1 \cdot s_0 \cdot x_{-\alpha}\left(-\epsilon bm' X + \tfrac{cmX}{1+am}\right) &\text{ by~\eqref{Chevalley-CCF1}.}
\end{align*}
Denote by $h_0$ the product of all factors in the above formula except the first and the last one. It is clear that $h_0$ lies in $P_\alpha(0)$. By~\cref{K-a-b} we can decompose $g$ as $g_0 \cdot x_{-\beta}(m'X) \cdot x_{-\alpha}(mX)$ for some $g_0 \in K(\alpha, \beta)$, $m, m' \in M$. The required assertion now follows from Lemmas~\ref{conj-S-a-x-bma}, \ref{lem:Smult} and the above computation:
\begin{multline} \nonumber
 x_{\beta - \alpha}(b) \cdot S_\alpha(a, g) \cdot h_{\alpha}^{-1}(1+am)\cdot h_{\beta - \alpha}(c^{-1})\cdot x_{\beta - \alpha}(-bc) = \\
  = x_{\beta - \alpha}(b) \cdot S_\alpha(a, g_0) \cdot x_{\beta-\alpha}(-b) \cdot h_0 \cdot x_{-\alpha}\left(\tfrac{m -\epsilon bm'}{1+am}\right) \in P_\alpha(\tfrac{m -\epsilon bm'}{1+am}). \qedhere
\end{multline}
\end{proof}

@@

\subsection{Proof of Horrocks theorem.} \label{sec:P1glueing}
We retain the notation of the previous section and assume additionally that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, R)$ is injective.
This assumption allows us to consider the group $\St(\Phi, B)$ as a subgroup of $\St(\Phi, R)$.
Being considered in this way, $\St(\Phi, B)$ contains the subgroup $\overline{G}^{\geq 0}_M$ of $\St(\Phi, R)$.
This follows from a consideration of the following commutative diagram, in which the map $t$ is obtained from~\cref{lem:lemma32}:
\[ \xymatrix{ \St(\Phi, A[X], M[X]) \ar[d] \ar[r] & \St(\Phi, R, I) \ar[r]^(.65){t} & \St(\Phi, B) \ar[d]^{j_R} \\
   \St(\Phi, A[X]) \ar[rr]^{j_+} & & \St(\Phi, R). } \]
